\subsection{Assumptions} \label{sec:limitations}

At a design level, we anticipate the schedule-relaxation approach to work
well for many programs/workloads as long as (1) they can benefit from
repeatability, (2) their schedules can be frequently reused, (3) their
races are rare, and (4) their nondeterminism comes from the sources
tracked by \peregrine.  This approach is certainly not designed for every
multithreaded program. For instance, like other DMT systems, \peregrine should
not be used for parallel simulators that desire nondeterminism for
statistical confidence.  For programs/workloads that rarely reuse
schedules, \peregrine may be unable to amortize the cost of recording and
analyzing execution traces.  For programs full of races, enforcing hybrid
schedules may be as slow as mem-schedules.  \peregrine addresses nondeterminism
due to thread scheduling and data races.  It mitigates input
nondeterminism by reusing schedules on different inputs.  It currently
considers command line arguments, data read from a file or a socket, and
the values returned by \v{random()}-variants as inputs.  \peregrine ensures that
schedule-reuses are fully deterministic if a program contains only these
nondeterminism sources, an assumption met by typical programs.  If a
program is nondeterministic due to other sources, such as functions that
query physical time (\eg, \v{gettimeofday()}), pointer addresses returned
by \v{malloc()}, and nondeterminism in the kernel or external libraries,
\peregrine relies on developers to annotate these sources.


%\peregrine leverages a set of techniques such as symbolic execution, program
% slicing, and alias analysis.  
The underlying techniques that \peregrine leverages make assumptions as well.
\peregrine computes preconditions from a trace slice using the symbolic
execution engine \klee, which does not handle floating point operations;
though recent work~\cite{klee-fp} has made advances in symbolic execution
of floating point programs.  (Note
that floating point operations not in trace slices are not an issue.)  We
explicitly designed \peregrine's slicing technique to compute sufficient
preconditions, but these preconditions may still include unnecessary ones,
because computing the \emph{weakest} (most relaxed) preconditions in general is
undecidable~\cite{aho:dragon:06}.  The alias analysis \peregrine uses makes a
few assumptions about the analyzed programs~\cite{alias:icse05};
a ``sounder'' alias analysis~\cite{alias:fse06} would remove these
assumptions.  These analyses may all get expensive for large
programs.  For server programs, \peregrine borrows the
windowing idea from our previous work~\cite{cui:tern:osdi10}; it is thus
similarly limited (\S\ref{sec:window}).

At an implementation level, \peregrine uses the LLVM framework, thus requiring
that a program is in either source (so we can compile using LLVM) or LLVM
IR format.  \peregrine ignores inline x86 assembly or calls to external
functions it does not know.  For soundness, developers have to lift x86
assembly to LLVM IR and provide summaries for external functions.  (The
external function problem is alleviated because \klee comes with a
\v{Libc} implementation.)  Currently \peregrine works only with a single process,
but previous work~\cite{dos:osdi10} has demonstrated how \dmt systems can
be extended to multiple processes.

%The preconditions \peregrine computes are not the \emph{weakest
%  preconditions}~\cite{aho:dragon:06} (\ie, most relaxed preconditions),


%relies on developers to annotate the boundaries of request processing and
%the functions that observe the internal state of the servers.

%% For instance, a parallel quick sort program may rarely reuse schedules
%% because it nondeterministically partitions data based on randomly chosen
%% pivots.  Fortunately, for performance, most parallel programs use
%% deterministic partitions of data, such as commonly used parallel sort with
%% sorting networks~\cite{sorting-networks} and 
%% all programs we have evaluated except \apache (because it does no
%% partitioning: it processes each client request with only one worker
%% thread).

% boundary: proxy process like dos

%% includes only the instructions dependent by the events
%% in the schedule.  control, data-dependency may not be widely
%% understood.  It does so by computing control- and data-dependencies,
%% within the same thread for reachability and across different threads for
%% determinism.  


